Predicting children's and adults' preferences in physical interactions via physics simulation

### 2021-05-10 paper/paper2021.Rmd
final edits

author-information: > 
    \author{
    {\large \bf George Kachergis}$^1,\star$, {\large \bf Samaher Radwan}$^1$, {\large \bf Bria Long}$^1$,  
    {\large \bf Judith Fan}$^2$, \\
    $^{\star}$\texttt{kachergis@stanford.edu} \\
    {\large \bf Daniel Bear}$^{1,3}$, {\large \bf Daniel Yamins}$^{1,3}$, {\large \bf Michael C.~Frank}$^1$ \\ 
    $^1$Department of Psychology, Stanford University \\
    $^2$Department of Psychology, University of California at San Diego \\
    $^3$Department of Computer Science, Stanford University}

@@

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
require(here)
require(tidyverse)
require(tidyboot)
library(apa) # for generating APA-style text to report statistical tests
library(ggpubr)
require(gridExtra)
require(kableExtra)
require(papaja)
```

@@

<!-- GK: possibly re-do this figure to have the proper orientations -->
```{r fig1-objects, fig.env = "figure", fig.pos = "h", fig.align='center', fig.width=3.5, fig.height=4.0, set.cap.width=T, num.cols.cap=1, fig.cap = "(Top) Nine of the 10 3D objects used for dropping and as targets (not pictured: ball). (Bottom) Example support and containment relations."}
img <- png::readPNG("figs/fig1_objects_support_contain_examples.png")
grid::grid.raster(img)
```

### Design
The experiment consisted of 20 drop trials, completed by each participant in one of four pseudorandom orders.
Participants were randomly assigned to each order.
Each drop trial displayed an object X that is to be dropped on one of two other objects (Y or Z), as described in the following procedure.
In the 20 trials, each of the above 3D objects was used twice as the drop object (X), and appeared four times as potential targets.
Instead of randomly sampling the target objects (Y or Z), each trial was designed to allow for one of two goal relations (or affordances): containment, or support.
That is, the target objects were chosen so that one of them would be capable of either supporting or  containing the dropped object.
For example, if asked to drop the cone on either the torus or the octahedron, a containment relation could be achieved between the cone and the torus, but not the octahedron (and support is unlikely due to the small surface area of the octahedron).
An example support relation is choosing to drop the pentagonal prism on the (equal diameter) pipe rather than the pyramid.
10 of the 20 trials were designed to afford containment relations, and the other 10 afforded support relations.
We refer to the objects affording the designed containment and support relations as *targets*, and the alternative objects as *distractors*.
Distractors for each trial were the same for all participants, and were selected to 1) not afford containment/support of the dropped object, and 2) roughly equate the frequency of appearance of all 10 objects.

### Procedure

Participants were instructed that they would be helping choose which toys to include in a new set of children's blocks in order to make them the most interesting.
Participants were then given a practice trial (see Figure 2), in which they were told that they should imagine dropping a given object (the torus) on each of two toys (the dumbbell or the pentagonal prism) in a bin. 
They were then asked to choose which toy to drop the torus on, and prompted to choose the most interesting or surprising combination.
After making their choice on the practice trial, participants were shown a 10-second video recording of 3D-printed plastic toys carrying out their chosen interaction (e.g., if they chose the dumbbell, they would see the torus dropped on the dumbbell).
This was done to ensure that they understood the consequences of the choice they made.
After the video of the torus drop, participants were given a sequence of 20 more drop trials asking them to choose which toy (X or Y) they would like to drop object Z on.
Four catch trials were interspersed among the 20 drop trials, which asked them to indicate which object they had just dropped on the previous trial (3-alternative forced choice).
The catch trials were meant to encourage attention to the objects, and participants who were incorrect on two or more of the catch trials were excluded.
<!-- Finally, there were two free response explanation trials mixed among the drop trials, querying participants, "Why did you make that choice?"
These free responses were examined to determine if participants had particular goals in mind when making a choice, and how prevalent those goals were among participants. -->

@@

combo_dat <- child_long %>% mutate(ID = as.character(ID),
                                Age = as.character(Age)) %>%
  select(ID, Age, Trial, choice, drop, target, relation, chose_target) %>%
  bind_rows(adult_long) %>% 
  mutate(AgeGroup = ifelse(Age!="adult", "child", "adult"))
#knitr::kable(trial_tab %>% arrange(relation, desc(adult_target_choice)), digits=2)

@@

## Results
We analyzed data from the `r Nad` adults who completed the experiment and answered at least 3 of the 4 catch trials correctly (45 participants were excluded for not meeting this criterion).
<!-- The main questions of interest are 1) are adults consistent in their preference of drop-target combinations? 2) are they more consistent with trials that afford support or containment relations? and 3) are there other apparent factors that drive their preferences? Our analysis addresses each question in turn. -->
Averaging each subject's binary responses (1=target relation, 0=alternative) for each trial type revealed a stronger preference for containment relations ($M=$ `r round(ad_rel_means[1,]$M,2)`) than support relations ($M=$ `r round(ad_rel_means[2,]$M,2)`; paired `r apa::t_apa(ad_supp_cont, format="rmarkdown", print=F)`).
Table 1 shows the proportion of participants that chose the designed target relation for each of the 20 trials. <!-- [should also show alternative target object?]. -->
Based on the binomial distribution, any of the trials on which more than 90 of the 155 participants agree (i.e., >0.59 or <0.41) significantly differ from chance. 
As can be seen in Table 1, adults participants (*adult* column) significantly preferred the target relation for all 10 of the containment trials, and significantly preferred five targets on the 10 support trials; the other five support trials did not significantly differ from chance, suggesting that many participants found the prospect of dropping on the distractor object at least equally enticing.

@@

<!-- Maybe code the type of goals stated by the parents? -->

@@

Overall, a few of these features were modestly correlated with people's choices. 
Table 2 shows the correlation between model preferences (with softmax $\beta = 1$) vs. adults' and children's choice proportions, both to all trials and separately for containment and support trials.
When all trials were considered together, the feature with the strongest correlation to adults' choices was the velocity of the target object (*SD(Target Vel)*, $r=.33$), while the feature with the strongest correlation to children's choices was *Support Sharpness* (see Methods, $r=.18$)
<!-- For support trials, both adults' ($r=.39$) and children's ($r=.41$) choices were best correlated with a feature indexing the variability in the target object's final position (*1/SD(Target Pos)*).  -->
<!-- For containment trials, adults' choices were most correlated with the velocity of the dropped object (*SD(Drop Obj Vel)*, $r=.54$) which also showed a modest correlation with children's choices ($r=.41$); however, children's choices were most correlated with the maximum distance of the target object (*Max(Target Dist)*, $r=.43$).  -->
Many of the other features showed weak or even moderate negative correlations with people's choices.

@@

At the same time, the fact that these results are not fully explainable by simple heuristic models suggest that this paradigm may fruitfully serve as a test bed for evaluating more complex computational theories of curiosity. 
For example, in one proposal learners prefer to explore stimuli that are "moderately discrepant" in relation to their current knowledge state, thereby providing an opportunity to learn [@Kinney1976], and curiosity changes as the gap between the learner's knowledge develops and the state of the world closes [@Loewenstein1994].
These theories are beginning to be implemented in deep neural networks, which are now capable of learning forward and inverse physical dynamics (i.e., "intuitive physics") from images when given the ability to "poke" the objects in the scene [@Agrawal2016], and have more recently been used to test which "curious" policies for generating actions result in robust and effective learning of intuitive physics in deep reinforcement learning agents [@Haber2018learning].
<!-- @Haber2018learning created a deep reinforcement learning model to drive intrinsically-motivated artificial agents, using two interacting neural networks: a *world-model* and a *self-model*.
These networks are trained simultaneously to solve different problems: the world-model attempts to predict the consequences of the agent's actions (either the forward or inverse dynamics), while the self-model attempts to learn to predict the errors of the world-model. 
The agent can use the predictions of the self-model to implement a variety of curiosity policies for choosing actions, for example to pick actions that it believes will challenge its world-model. 
-->
In future work, we aim to examine the development of curiosity in such curious artificial agents, and compare whether these agents behave similarly to people when asked to "drop it like it's hot" [@Snoop].

The multifaceted nature of curiosity has fascinated researchers for decades, but only recently have we developed the means to factorize and test these theories. 
We propose that the development of parallel tasks for children, humans, and embodied agents is a promising way forward towards modeling curiosity in the real-world.

<!--According to these theories, curiosity-driven behavior reflects a gap between the learner's knowledge and the state of the world [@Loewenstein1994], and thus they predict that as a learner gains additional knowledge, their preferences will shift toward more complex stimuli [@Dember1957]
Although these theories have provided important qualitative insights, one of their major limitations is that they do not provide precise ways to characterize a learner's current knowledge state, how physical states of the world are represented in their mind, nor how discrepancies between a learner's knowledge and such states are compared.  
Our approach leverages theory and insights from reinforcement learning (RL), including the value of mechanisms that equip RL agents with "intrinsic motivation" — a drive to explore the state space even when rewards are sparse or absent, and favor actions with uncertain outcomes, because they may result in the discovery of new policies with high expected values [@Schmidhuber2010].
Our modeling approach is resonant with recent work that has instantiated such intrinsic-motivation mechanisms in robots to help them learn robust ways to predict physical events in the world world [@Oudeyer2007; @Oudeyer2013], although this work typically involves pretraining models on a separate physical prediction task before implementing curiosity-driven learning, and has not yet been directly compared to human behavior in the same prediction tasks.
-->
<!--
Is containment inherently more satisfying, or is there greater interest because it is more achievable? (Less susceptible to small amounts of error)
Are children less interested in relations that are more difficult to achieve with more motor noise?
Or are they less aware (or simply less interested in) support?
Or do children find some other goal relatively more enticing?
-->

>>>>

author-information: > 
    \author{
    {\large \bf George Kachergis}$^{1,\star}$, {\large \bf Samaher Radwan}$^1$, {\large \bf Bria Long}$^1$,  
    {\large \bf Judith Fan}$^2$, \\
    $^{\star}$\texttt{kachergis@stanford.edu} \\
    {\large \bf Daniel Bear}$^{1,3}$, {\large \bf Daniel Yamins}$^{1,3}$, {\large \bf Michael C.~Frank}$^1$ \\ 
    $^1$Department of Psychology, Stanford University \\
    $^2$Department of Psychology, University of California at San Diego \\
    $^3$Department of Computer Science, Stanford University}

@@

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
require(here)
require(tidyverse)
require(tidyboot)
library(apa) # for generating APA-style text to report statistical tests
library(ggpubr)
require(gridExtra)
require(papaja)
require(kableExtra)
```

@@

```{r fig1-objects, fig.env = "figure", fig.pos = "h", fig.align='center', fig.width=3.5, fig.height=4.0, set.cap.width=T, num.cols.cap=1, fig.cap = "(Top) Nine of the 10 3D objects used for dropping and as targets (not pictured: ball). (Bottom) Example support and containment relations."}
img <- png::readPNG("figs/fig1_objects_support_contain_examples.png")
grid::grid.raster(img)
```

### Design
The experiment consisted of 20 drop trials, completed by each participant in one of four pseudorandom orders.
Participants were randomly assigned to each order.
Each drop trial displayed an object X that is to be dropped on one of two other objects (Y or Z), as described in the following procedure.
In the 20 trials, each of the above 3D objects was used twice as the drop object (X), and appeared four times as potential targets.
Instead of randomly sampling the target objects (Y or Z), each trial was designed to allow for one of two goal relations (or affordances): containment, or support.
That is, the target objects were chosen so that one of them would be capable of either supporting or  containing the dropped object (see examples: Figure 1, bottom).
For example, if asked to drop the cone on either the torus or the octahedron, a containment relation could be achieved between the cone and the torus, but not the octahedron (and support is unlikely due to the small surface area of the octahedron).
An example support relation is choosing to drop the pentagonal prism on the (equal diameter) pipe rather than the pyramid.
10 of the 20 trials were designed to afford containment relations, and the other 10 afforded support relations.
We refer to the objects affording the designed containment and support relations as *targets*, and the alternative objects as *distractors*.
Distractors for each trial were the same for all participants, and were selected to 1) not afford containment/support of the dropped object, and 2) roughly equate the frequency of appearance of all 10 objects.

### Procedure
Participants were instructed that they would be helping choose which toys to include in a new set of children's blocks in order to make them the most interesting.
Participants were then given a practice trial (see Figure 2), in which they were told that they should imagine dropping a given object (the torus) on each of two toys (the dumbbell or the pentagonal prism) in a bin. 
They were then asked to choose which toy to drop the torus on, and prompted to choose the most interesting or surprising combination.
After making their choice on the practice trial, participants were shown a 10-second video recording of 3D-printed plastic toys carrying out their chosen interaction (e.g., if they chose the dumbbell, they would see the torus dropped on the dumbbell).
This was done to ensure that they understood the consequences of the choice they made.
After the video of the torus drop, participants were given a sequence of 20 more drop trials asking them to choose which toy (X or Y) they would like to drop object Z on.
Four catch trials were interspersed among the 20 drop trials, which asked them to indicate which object they had just dropped on the previous trial (3-alternative forced choice).
The catch trials were meant to encourage attention to the objects, and participants who were incorrect on two or more of the catch trials were excluded.
<!-- Finally, there were two free response explanation trials mixed among the drop trials, querying participants, "Why did you make that choice?"
These free responses were examined to determine if participants had particular goals in mind when making a choice, and how prevalent those goals were among participants. -->

@@

combo_dat <- child_long %>% mutate(ID = as.character(ID),
                                Age = as.character(Age)) %>%
  select(ID, Age, Trial, choice, drop, target, relation, chose_target) %>%
  bind_rows(adult_long) %>% 
  mutate(AgeGroup = ifelse(Age!="adult", "child", "adult")) %>%
  mutate(Age = factor(Age, levels=c("3","4","5","6","7","adult")))
#knitr::kable(trial_tab %>% arrange(relation, desc(adult_target_choice)), digits=2)

@@

## Results
We analyzed data from the `r Nad` adults who completed the experiment and answered at least 3 of the 4 catch trials correctly (45 participants were excluded for not meeting this criterion).
<!-- The main questions of interest are 1) are adults consistent in their preference of drop-target combinations? 2) are they more consistent with trials that afford support or containment relations? and 3) are there other apparent factors that drive their preferences? Our analysis addresses each question in turn. -->
Averaging each subject's binary responses (1=target relation, 0=alternative) for each trial type revealed a stronger preference for containment relations ($M=$ `r round(ad_rel_means[1,]$M,2)`) than support relations ($M=$ `r round(ad_rel_means[2,]$M,2)`; paired `r apa::t_apa(ad_supp_cont, format="rmarkdown", print=F)`).
Table 1 shows the proportion of participants that chose the designed target relation for each of the 20 trials. <!-- [should also show alternative target object?]. -->
Based on the binomial distribution, any of the trials on which more than 90 of the 155 participants agree (i.e., >0.59 or <0.41) significantly differ from chance. 
As can be seen in Table 1, adult participants significantly preferred the target relation for all 10 of the containment trials, and significantly preferred five targets on the 10 support trials; the other five support trials did not significantly differ from chance, suggesting that many participants found the prospect of dropping on the distractor object at least equally enticing.

@@

```{r regression, eval=F, include=F}
require(lme4)
m1 <- lmer(chose_target ~ AgeGroup * relation + (1|ID) + (1|Trial), data=combo_dat)
summary(m1) # children less likely to choose target than adults, and support targets less likely picked than containment targets
```

@@

Overall, a few of these features were modestly correlated with people's choices. 
Table 2 shows the correlation between model preferences (with softmax $\beta = 1$) vs. adults' and children's choice proportions, both to all trials and separately for containment and support trials.
When all trials were considered together, the feature with the strongest correlation to adults' choices was the velocity of the target object (*SD(Target Vel)*, $r=.33$), while the feature with the strongest correlation to children's choices was *Support Sharpness* (see Methods, $r=.18$).
For support trials, both adults' ($r=.39$) and children's ($r=.41$) choices were best correlated with a feature indexing the variability in the target object's final position (*1/SD(Target Pos)*).
For containment trials, adults' choices were most correlated with the velocity of the dropped object (*SD(Drop Obj Vel)*, $r=.54$) which also showed a modest correlation with children's choices ($r=.41$); however, children's choices were most correlated with the maximum distance of the target object (*Max(Target Dist)*, $r=.43$).
Many of the other features showed weak or even moderate negative correlations with people's choices.

@@

At the same time, the fact that these results are not fully explainable by simple heuristic models suggest that this paradigm may fruitfully serve as a test bed for evaluating more complex computational theories of curiosity. 
For example, in one proposal learners prefer to explore stimuli that are "moderately discrepant" in relation to their current knowledge state, thereby providing an opportunity to learn [@Kinney1976], and curiosity changes as the gap between the learner's knowledge develops and the state of the world closes [@Loewenstein1994].
These theories are beginning to be implemented in deep neural networks, which are now capable of learning forward and inverse physical dynamics (i.e., "intuitive physics") from images when given the ability to "poke" the objects in the scene [@Agrawal2016], and have more recently been used to test which "curious" policies for generating actions result in robust and effective learning of intuitive physics in deep reinforcement learning agents [@Haber2018learning].

Such embodied deep reinforcement learning agents are often composed of two models: a *world model* which attempts to predict the consequences of the agent's actions (i.e., the forward dynamics--physics--of the world), while the *self model* attempts to predict the errors of the world model.
The predictions of the self model can be used to implement different curiosity policies for choosing actions [e.g., picking actions that are expected to challenge its world model; @Haber2018learning].
In future work, we aim to examine the development of curiosity in such curious artificial agents, and compare whether these agents behave similarly to people when asked to "drop it like it's hot" [@Snoop].
Will these agents show choices consistent with potential containment/support relations, or do these human-like preferences require representation and selection of possible goals?
How might the agents' choices evolve over time, as the world model matures?
Might the agents' curiosity policies need to change over development in order to match people's choices, or do these goals emerge through the dynamics of learning?

The multifaceted nature of curiosity has fascinated researchers for decades, but only recently have we developed the means to factorize and test these theories. 
We propose that the development of parallel tasks for children, humans, and embodied agents is a promising way forward towards modeling curiosity in the real world.

<!-- Is containment inherently more satisfying, or is there greater interest because it is more achievable? (Less susceptible to small amounts of error)
Are children less interested in relations that are more difficult to achieve with more motor noise?
Or are they less aware (or simply less interested in) support?
Or do children find some other goal relatively more enticing? -->

<!--According to these theories, curiosity-driven behavior reflects a gap between the learner's knowledge and the state of the world [@Loewenstein1994], and thus they predict that as a learner gains additional knowledge, their preferences will shift toward more complex stimuli [@Dember1957]
Although these theories have provided important qualitative insights, one of their major limitations is that they do not provide precise ways to characterize a learner's current knowledge state, how physical states of the world are represented in their mind, nor how discrepancies between a learner's knowledge and such states are compared.  
Our approach leverages theory and insights from reinforcement learning (RL), including the value of mechanisms that equip RL agents with "intrinsic motivation" — a drive to explore the state space even when rewards are sparse or absent, and favor actions with uncertain outcomes, because they may result in the discovery of new policies with high expected values [@Schmidhuber2010].
Our modeling approach is resonant with recent work that has instantiated such intrinsic-motivation mechanisms in robots to help them learn robust ways to predict physical events in the world world [@Oudeyer2007; @Oudeyer2013], although this work typically involves pretraining models on a separate physical prediction task before implementing curiosity-driven learning, and has not yet been directly compared to human behavior in the same prediction tasks.
-->
